---
title: "LMS Statistical Extracts"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This is a repository to help researchers use statistical extracts produced as part of two papers coauthored by Lamadon, Mogstad, and Setzler:

- "Imperfect competition, compensating differentials, and rent sharing in the US labor market" by Thibaut Lamadon, Magne Mogstad, and Bradley Setzler. American Economic Review 112.1 (2022): 169-212.

- "How Much Should we Trust Estimates of Firm Effects and Worker Sorting?" by Stéphane Bonhomme, Kerstin Holzheu, Thibaut Lamadon, Elena Manresa, Magne Mogstad, and Bradley Setzler. Working Paper w27368. National Bureau of Economic Research (2021).





## Value Added and Wage Growth for Stayers in the US Tax Data

Here, we provide the estimates of the regional pass-through from firm productivity shocks to the earnings of workers.  In particular, we provide the region-specific changes in log earnings and log value added relative to event time zero at which a log value added shock occurs using the   approach described under the heading ``Difference-in-Difference Illustration of Internal Instruments" in subsection III.A of [LMS](https://www.bradleysetzler.com/files/Lamadon-Mogstad-Setzler.pdf). The provided variables are:

- `EventTime`: This is the number of years since the treated group received a positive log value added shock and the control group did not.
- `LogVADiff_Unconditional` and `LogWageDiff_Unconditional`: These variables are the mean difference in firm-specific log value added and log earnings among stayers, respectively, between the treated group and control group at a particular event time. 
- `LogVADiff_MarketLevel` and `LogWageDiff_MarketLevel`: These variables are the mean difference in market-level averages of log value added and log earnings among stayers, respectively, between the treated group and control group at a particular event time. 
- `LogVADiff_FirmLevel` and `LogWageDiff_FirmLevel`:  These variables are similar to `LogVADiff_Unconditional` and `LogWageDiff_Unconditional`, respectively, but controlling for market-year fixed effects.

The statistical extract containing these estimates is available for download [here](https://raw.githubusercontent.com/setzler/LMS-Statistical-Extracts/main/StatisticalExtracts/LMS-Stayer-Differences.csv). The first 3 rows of the statistical extract are as follows:

``` {r echo = FALSE, results = 'asis'}
library(data.table)
dd = setDT(read.csv(file="StatisticalExtracts/LMS-Stayer-Differences.csv"))
knitr::kable(dd[1:3])
```






## Region-Sector Pass-through Measures for Stayers in the US Tax Data

Here, we provide the region-specific instrumental variables estimation of the pass-through from log value added to  log earnings using the approach described under the heading ``Formal Identification Using Internal Instruments" in subsection III.A of [LMS](https://www.bradleysetzler.com/files/Lamadon-Mogstad-Setzler.pdf).  Denoting log earnings by $w$ and log value added by $y$, the empirical equation is,

$$ \mathbb{E}[ \Delta y_t (w_{t+e} - w_{t-e'} - \gamma_r(y_{t+e} - y_{t-e'}))  ] = 0 $$
which is estimated only among workers who are stayers and which is estimated separately for each broad market $r$. A similar equation is estimated using market-level means across all broad markets, in which case the parameter estimate is denoted $\Upsilon$.The preferred estimates of $\gamma_r$ are based on firm-level deviations from market-level means.  The estimated $(\gamma_r,\Upsilon)$ parameters can then be transformed into the preference parameters $(\rho_r,\beta)$.

The statistical extract contains the following estimates:

- `BroadMarket`: This variable indicates the broad market, defined as a region and a sector of the economy.
- `UnconditionalPassthrough`: This variable is the IV estimate of $\gamma_r$ that does not control for market-year fixed effects.
- `NetPassthrough`: This variable is the preferred IV estimate of $\gamma_r$ that  controls for market-year fixed effects.
- `MarketPassthrough`: This variable is the IV estimate of $\Upsilon$ using only market-level aggregate data; this estimate does not vary across broad markets.
- `Parameter_rho`: This variable is the estimate of the within-market preference parameter $\rho_r$, which varies across broad markets.
- `Parameter_beta`: This variable is the estimate of the between-market preference parameter $\beta$; this estimate does not vary across broad markets.

Estimating this equation requires population employer-employee data on earnings and firm value added. We provide estimates of the various terms in this equation. The statistical extract containing these estimates is available for download [here](https://raw.githubusercontent.com/setzler/LMS-Statistical-Extracts/main/StatisticalExtracts/LMS-Regional-Passthrough.csv). The first 3 rows of the statistical extract are as follows:

``` {r echo = FALSE, results = 'asis'}
library(data.table)
dd = setDT(read.csv(file="StatisticalExtracts/LMS-Regional-Passthrough.csv"))
knitr::kable(dd[1:3])
```






## Region-Sector Rents  in the US Tax Data

Here, we provide the region-specific rent estimation using the approach of Result 4 from subsection II.E of [LMS](https://www.bradleysetzler.com/files/Lamadon-Mogstad-Setzler.pdf). In particular, equation (9) of LMS shows that the firm technology parameter $\alpha_r$ can be recovered for each market using only the parameters $(\rho_r,\beta)$ discussed above and the (quasi) labor share. Then, given parameters $(\alpha_r,\rho_r,\beta)$, Result 4 of LMS shows that per-capita rents can be recovered from the per-capita wagebill and per-capita value added.

The statistical extract contains the following estimates:

- `BroadMarket`: This variable indicates the broad market, defined as a region and a sector of the economy.
- `SampleWorkerYears`: This variable is the count of worker-year observations used in the estimation, measured in 1,000 units.
- `SampleFirmYears`: This variable is the count of firm-year observations used in the estimation, measured in 1,000 units.
- `LaborShare`: This variable is the estimate of the (quasi) labor share defined as $\exp( \mathbb{E}[\ell + w - y])$, where $\ell$ is log labor, $w$ is log earnings, and $y$ is log value added.
- `ValueAdded`: This variable is the estimate of average value added, in per-capita units.
- `Wagebill`: This variable is the estimate of average wage bill, in per-capita units.
- `Parameter_alpha`: This variable is the estimate of the market-specific firm productivity parameter $\alpha_r$.
- `WorkerRents_FirmLevel` and `FirmRents_FirmLevel`: These variables are the estimates of rents at the firm-level that are captured by workers and by firms, respectively, measured as $1  per-capita.
- `WorkerRents_MarketLevel` and `FirmRents_MarketLevel`: These variables are the estimates of rents at the market-level that are captured by workers and by firms, respectively, measured as $1  per-capita.

We provide estimates of the various terms in this equation. The statistical extract containing these estimates is available for download [here](https://raw.githubusercontent.com/setzler/LMS-Statistical-Extracts/main/StatisticalExtracts/LMS-Regional-Rents.csv). The first 3 rows of the statistical extract are as follows:

``` {r echo = FALSE, results = 'asis'}
library(data.table)
dd = setDT(read.csv(file="StatisticalExtracts/LMS-Regional-Rents.csv"))
knitr::kable(dd[1:3])
```






## Earnings Inequality and Firm-Worker Clusters in the US Tax Data

Here, we provide the estimates of the equilibrium earnings expression presented in equation (14) of [LMS](https://www.bradleysetzler.com/files/Lamadon-Mogstad-Setzler.pdf) using the empirical specification and estimation approach described in subsection IV.A of LMS. The empirical equation is,

$$w^a_{k}(x) = \psi_k + x\theta_k $$

where $k$ denotes a cluster of firms, $x$ denotes the measure of worker skill, $w^a$ is the predicted log wage after residualization and adjustment for productivity shocks, $\psi$ is the cluster-specific wage premium, and $\theta$ is the cluster-specific skill complementarity.

Estimating this equation requires population employer-employee data on earnings and firm value added. We provide estimates of the various terms in this equation, evaluated at quantiles $q$ in the $x$ distribution. In particular, the variables are:

- `FirmCluster_k`: This variable indicates the cluster $k$ of firms.
- `Size_FirmCluster_k`: This variable indicates the number of workers employed by cluster $k$.
- `FirmPremium_psik`: This variable indicates the firm premium $\psi_k$ for cluster $k$.
- `SkillComplementarity_thetak`: This variable indicates the skill complementarity $\theta_k$ for cluster $k$.
- `WorkerDecile_q`: This variable indicates quantile $q$ in the distribution of $x$.
- `WorkerSkill_xq`: This variable indicates the value of worker skill $x$ at quantile $q$.
- `ExpectedLogWage_wkq`: This variable indicates the expected log wage $w^a$ for cluster $k$ and worker skill quantile $q$, which is residualized on observables and adjusted for productivity shocks.

The statistics are available for download [here](https://raw.githubusercontent.com/setzler/LMS-Statistical-Extracts/main/StatisticalExtracts/LMS-Firm-Worker-Matches.csv).

``` {r echo = FALSE, results = 'asis'}
dd = setDT(read.csv(file="StatisticalExtracts/LMS-Firm-Worker-Matches.csv"))
knitr::kable(dd[1:3])
```

In case the user prefers not to use the LMS adjustment for productivity shocks, the unadjusted version of these estimates are available for download  [here](https://raw.githubusercontent.com/setzler/LMS-Statistical-Extracts/main/StatisticalExtracts/LMS-Firm-Worker-Matches_unadjusted.csv).








## Between and Within Firm Inequality in Tax Data from the US and 4 European Countries

[BHLMMS](https://www.bradleysetzler.com/files/BHLMM-Setzler.pdf) analyze employer-employee administrative data from 5 countries: Austria, Italy, Norway, Sweden, and the US. They take extensive steps to harmonize the various datasets to ensure cross-country comparability.

They characterize the role of firms in wage inequality by performing a within-between decomposition. Letting $w$ denote the log wage (or earnings), and letting $j(it)$ denote the firm at which worker $i$ is employed in year $t$, the average wage in firm $j$ is given by $w_j \equiv \mathbb{E}[w_{it} | j(it)=j]$. Then, we can write $w_{it} = w_j + (w_{it} - w_j)$, so the total variance in log wages can be expressed,
$$\text{Var}(w_{it}) = \text{Var}(w_{j}) + \text{Var}(w_{it} - w_j)$$
where $\text{Var}(w_{j})$ is the between-firm variance and $\text{Var}(w_{it} - w_j)$ is the within-firm variance.

In addition to providing the within-between decomposition, [BHLMMS](https://www.bradleysetzler.com/files/BHLMM-Setzler.pdf) describe the connectedness of each country's firms by examining movers across firms and analyzing how many firms are workers are contained the full data, the connected subset of firms, and the leave-one-out connected subset of firms. The variables provided are:

- `Country`: This variable indicates the country represented by the data.
- `YearRange`: This variable indicates the number of years included in the sampling timeframe.
- `Set`: The set can be all firms, the connected subset of firms, or the leave-one-out connected subset of firms
- `WorkerCount`: This variable is the count of workers used in the estimation.
- `MoverCount`: This variable is the count of movers (i.e. workers who switch employers) used in the estimation.
- `FirmCount`: This variable is the count of firms used in the estimation.
- `LogWageVariance`: This is the variance in log wages (or earnings), $\text{Var}(w_{it})$.
- `BetweenFirmVariance`: This is the between-firm variance in log wages (or earnings), $\text{Var}(w_{j})$.

The statistics are available for download [here](https://raw.githubusercontent.com/setzler/LMS-Statistical-Extracts/main/StatisticalExtracts/LMS-Firm-Worker-Matches.csv).  The first 3 rows of the statistical extract are as follows:

``` {r echo = FALSE, results = 'asis'}
dd = setDT(read.csv(file="StatisticalExtracts/BHLMMS-WithinBetween-Variance.csv"))
knitr::kable(dd[1:3])
```


## Firm Effects and Sorting in Tax Data from the US and 4 European Countries


Here, we provide the estimates of the two-way fixed-effects decomposition from equation (3) of [BHLMMS](https://www.bradleysetzler.com/files/BHLMM-Setzler.pdf) using 4 estimators applied to data from 5 countries. Assuming that the residual  log wage $w$ can be expressed as $w_{it} = x_i + \psi_{j(it)} + \epsilon_{it}$, where $j(it)$ denotes the firm at which worker $i$ is employed in year $t$, then the variation in the log wage can be expressed,

$$\text{Var}(w_{it}) = \text{Var}(x_i) +\text{Var}(\psi_{j(it)}) + 2\text{Cov}(\psi_{j(it)},x_i) +  \text{Var}(\epsilon_{it})$$
The two key components of interest are $\text{Var}(\psi_{j(it)})$, which is the variance in log wages due to firm premia, and $2\text{Cov}(\psi_{j(it)},x_i)$, which is the variance in log wages due to the sorting of worker skill to firm premia. 

Estimating this equation requires population employer-employee data on earnings from each of the countries. The 4 estimation methods are FE, CRE, FE-HO, FE-HE; see Section 4 of [BHLMMS](https://www.bradleysetzler.com/files/BHLMM-Setzler.pdf) for discussion of the methods. The variables are:

- `Country`: This variable indicates the country represented by the data.
- `YearRange`: This variable indicates the number of years included in the sampling timeframe.
- `Estimator`: This variable indicates which of the methods was used in the estimation.
- `LogWageVariance`: This is the variance in log wages (or earnings), $\text{Var}(w)$.
- `FirmEffectVariance`: This is the variance in the firm effects, $\text{Var}(\psi)$.
- `FirmWorkerEffectsCovariance`: This is the covariance between the firm effects and the worker effects, $\text{Cov}( \psi , x )$.


The estimates using each of those methods are available for download [here](https://raw.githubusercontent.com/setzler/LMS-Statistical-Extracts/main/StatisticalExtracts/BHLMMS-FirmEffects-Sorting.csv).  The first 3 rows of the statistical extract are as follows:

``` {r echo = FALSE, results = 'asis'}
dd = setDT(read.csv(file="StatisticalExtracts/BHLMMS-FirmEffects-Sorting.csv"))
knitr::kable(dd[1:3])
```




